---
title: "3D Linear Regression Surfaces with Marginal Effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3D Linear Regression Surfaces with Marginal Effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = F
)
```


## Introduction

Multiple regressions create regression **surfaces**, which have as many dimensions as there are variables (both $x$ and $y$) in the multiple regression equation. Marginal effects are a direction in this surface and can be visualized by a line in the surface. 

This vignette demonstrates a 3 dimensional visualization of marginal effects (`add_marginals`) and regression surface (`add_3d_surface()`) for a linear regression with two $x$ variables. This kind of model creates a flat regression surface.

Other types of regression surfaces are shown in `vignette("linear_models_w_interactions_3d")` and `vignette(generalized_linear_models_3d)`.

## Setup 

The package __regress3d__ is built on the syntax of __plotly__ in R. Both libraries are called in the initial setup.  

```{r setup, message = F, warning=F}
library(regress3d)
library(plotly)
```

## Linear Regression Surfaces

The basic functionality of the functions `add_3d_surface()` and `add_marginals()` are demonstrated in this vignette. Both functions:

* default to the plot data specified in the call to `plot_ly()`,
* have an optional `ci` argument to visualize the confidence intervals, which defaults to `TRUE`.

Both the marginal effects and the regression surface shown below, but either can be plotted independently of the other.


```{r, out.width = "800px", out.height = "450px"}

mymodel <- lm(r_shift ~ median_income16 + any_college, 
              data = county_data, weight = pop_estimate16)

plot_ly( data = county_data,
         x = ~median_income16,
         y = ~any_college,
         z = ~r_shift) %>%
  add_3d_surface(model = mymodel)%>%
  add_marginals(model = mymodel)
```


-   Click and drag on the graphic to rotate it.
-   When needed, you can reset the image to its starting point. Hover over the image and click on the house symbol at the top right above the graphic title. Alternately, you can reload the page.
-   The regression surface is shown in the flat dark blue surface. 
-   The gray curved surfaces above and below the regression surface represent the 95\% confidence intervals.
-   The marginal effects and their corresponding confidence intervals are shown in [red]{style="color: crimson;"} and [orange]{style="color: orange;"} lines. 

## Improvements with Plotly

We can make this interactive image prettier by adding scatterpoints with the __plotly__ function `add_markers()`, a hovertemplate for each point, and better labels. The library __scales__ helps present the numeric values, percentages, and dollars more clearly.

```{r, out.width = "800px", out.height = "450px"}
library(scales)
plot_ly( data = county_data,
         x = ~median_income16,
         y = ~any_college,
         z = ~r_shift) %>%
  add_markers(size = ~pop_estimate16, 
              hovertemplate = ~paste("<b>", county_state,"</b><br>", 
                            "County population: ",comma(pop_estimate16),"<br><br>",
                            "County median income: ",dollar(median_income16),"<br>", 
                            "County college experience: ",percent(any_college, 
                                                                  accuracy = 0.1, 
                                                                  scale = 1),"<br>", 
                            "Shift to Trump: ",percent(r_shift, 
                                                       accuracy = 0.01, scale = 1),"<br>"),
              showlegend = F, name = "county",
              color = I("black")) %>%
  add_3d_surface(model = mymodel) %>%
  add_marginals(model = mymodel) %>%
  layout( 
    title = "\nCounty shift to Trump in 2016",
    scene = list(xaxis = list(title = 'County median income'),
                 yaxis = list(title = 'County college experience'),
                 zaxis = list(title = 'Shift to Trump')),
    legend = list(font = list(size = 8))
    )
```
