---
title: "Interaction Terms in 3D"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interaction Terms in 3D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = F,
  message = F
)
```


## Introduction
This vignette examines the functions `add_3d_surface()` and `add_marginals()` for a linear regression surface that features an interaction term. The functionality of `add_3d_surface()` and `add_marginals()` is the same as in a linear regression model without an interaction term. However, the visuals are sufficiently different to merit a separate description.

## Setup 

The package __regress3d__ is built on the syntax of __plotly__ in R. Both libraries are called in the initial setup.  

```{r setup, message = F, warning=F}
library(regress3d)
library(plotly)
```

## Interaction Terms

An interaction term in a linear regression can be depicted as a curved regression surface with linear marginal effects. The marginal effects of $x_1$ ($x_2$) are allowed to linearly change as $x_2$ ($x_1$) changes. This captures heterogenous effects in a regression, where the effect of $x_1$ is different for different values of $x_2$. Note that the change in the effect of $x_1$ as $x_2$ changes is restricted by the linearity of the interaction term.

The equation for the predicted values from a linear regression with two $x$ variables and an interaction term is below.

\begin{align*}
\hat{y} =& \hat{\beta}_1 x_1 +\hat{\beta}_2 x_2 +\hat{\beta}_{12} x_1 x_2 +  \hat{\beta}_0 
\end{align*}

### Data

The variables in this example are county level demographic and voting measures from 2016. The data was chosen to emphasize the twist in the regression surface and highlight change in the marginal effects. The explanatory variable `college_2cat_num` was chosen to be binary to make it easier to see how the marginal effect for one value of the binary variable is different from the marginal effect for the other value.

* `prcnt_GOP16`: the percent of the county's voters who voted for Trump in 2016,
* `prcnt_unemployed_log`: the logged percent of the county that was unemployed in 2016, and 
* `college_2cat_num`: a binary measure of county college experience. It is 0 if the county was below the median value of adult residents who had been enrolled in college at some point regardless of whether they graduated, and 1 otherwise.

The regression is weighted by `pop_estimate16`, the number of people in a county, to capture the influence of large counties.

### Code and graphic

The code starts by specifying a model. We then create a `plotly::plot_ly()` object using the same variables. Next layer on the scattercloud, 3D surface, marginals, and labels. Note that regression notation uses $x_1$ and $x_2$ to represent the explanatory variables, and $y$ for the outcome. The plotly command denotes the explanatory variables as $x$ and $y$, and the outcome variable is $z$.

```{r, out.width = "800px", out.height = "450px"}

county_data$prcnt_unemployed_log <- log(county_data$prcnt_unemployed)

mymodel <- lm(prcnt_GOP16~ prcnt_unemployed_log * college_2cat_num, 
              data = county_data, weights = pop_estimate16)

plot_ly( data = county_data,
         x = ~prcnt_unemployed_log,
         y = ~college_2cat_num,
         z = ~prcnt_GOP16) %>%
  add_3d_surface(model = mymodel)%>%
  add_marginals(model = mymodel, x1_constant_val = 2.5, x2_constant_val =0)%>%
  add_marginals(model = mymodel, x1_constant_val = 1, x2_constant_val =1) 
```

See the article on [linear models](https://ellafostermolina.github.io/regress3d/articles/linear_models_3d.html#code-and-graphic) for more information on how to manipulate this graphic. 

The marginal effects are called twice in this example to highlight two key features of the regression surface. 

* The [orange]{style="color: orange;"} marginal effects of $x_1$ are linear for any given value of $x_2$. The same is true for the [red]{style="color: red;"} marginal effects of $x_2$.
* The size of the marginal effect of $x_1$ ($x_2$) is different for the two values of $x_2$ ($x_1$) plotted.

## More on marginal effects

Because $x_2$ (education) is binary, the magnitude of the marginal effect of $x_1$ (unemployment) when $x_2$ has the value of $0$ and $1$ is easily recoverable. Simply estimate the regression on the subset of data where education has the value of 1, and again for 0. The standard errors for those subsets are different from those depicted in the image, but the magnitude of the effects are the same. 

We can highlight this fact in the image using __plotly__, __scales__, and __regress3d__.

Using __plotly__ functions and arguments, we add scatterpoints, a hovertemplate for each point, and better labels. 

The library __scales__ helps present the numeric values and percentages in the hovertext more clearly.

__regress3d__ lets us modify the legend text by changing `x1_direction_name` and `x2_direction_name`. The default is "Predicted marginal effect of x1". It is possible, but not shown here, to adjust which value $x_2$ ($x_1$) is held constant at. 

```{r, out.width = "800px", out.height = "450px"}
library(scales)

county_data$prcnt_unemployed_log <- log(county_data$prcnt_unemployed)
mymodel <- lm(prcnt_GOP16~ prcnt_unemployed_log * college_2cat_num, 
              data = county_data, weights = pop_estimate16)


plot_ly(data = county_data, 
              x = ~prcnt_unemployed_log,
              y = ~college_2cat_num,
              z = ~prcnt_GOP16 ) %>%
  add_markers(size = ~pop_estimate16,
              hovertemplate = ~paste("<b>", county_state,"</b><br>",
                            "County population: ",comma(pop_estimate16),
                            "<br><br>", 
                            "County unemployment: ",percent(prcnt_unemployed, accuracy = 0.1, scale = 1),"<br>",
                            "County college experience: ",
                            percent(any_college, accuracy = 0.1, scale = 1),"<br>",
                            "Trump Vote %: ",
                            percent(prcnt_GOP16, accuracy = 0.01, scale = 1),"<br>"),
              showlegend = F, name = "county",
              color = ~college_2cat,
              colors = c("deepskyblue", "dimgrey"))  %>%
  add_3d_surface(model = mymodel)%>%
  add_marginals(model = mymodel, x1_constant_val = 2.5, x2_constant_val =0,
                x1_color = "yellow",
                x2_color = "grey",
                x1_direction_name = "Marginal effect of unemployment<br>for low education counties",
                x2_direction_name = "")%>%
  add_marginals(model = mymodel, x1_constant_val = 1, x2_constant_val =1,
                x2_color = "grey",
                x1_direction_name = "Marginal effect of unemployment<br>for high education counties",
                x2_direction_name = "") %>%
  layout( 
    title = "\nCounty Vote for Trump by\n unemployment & education",
    scene = list(xaxis = list(title = 'logged unemployment'),
                 yaxis = list(title = 'education'),
                 zaxis = list(title = '% vote for Trump 2016')),
    # width = 500, height = 500,
    legend = list(font = list(size = 8))
    )
```

# More on marginal effects

# Numeric regression results

These graphics correspond to the following numerical regression results. Three models are presented to allow the reader to examine the 

1.  The multiple regression equation with an interaction term.
2.  A simple regression using only education. 

```{r, results='asis', echo =F}
library(stargazer)

county_data$prcnt_unemployed_log <- log(county_data$prcnt_unemployed)

unemployed_model_edu0 <- lm(prcnt_GOP16~ prcnt_unemployed_log, 
              data = subset(county_data, college_2cat_num == 0), 
              weights = pop_estimate16)
unemployed_model_edu1 <- lm(prcnt_GOP16~ prcnt_unemployed_log, 
              data = subset(county_data, college_2cat_num == 1), 
              weights = pop_estimate16)
two_vars_model <- lm(prcnt_GOP16~ prcnt_unemployed_log * college_2cat_num, 
              data = county_data, weights = pop_estimate16)

stargazer(two_vars_model, unemployed_model_edu0, unemployed_model_edu1,
          type = 'html',
          keep.stat = c("n", "adj.rsq"),
          star.cutoffs = c(0.05),
          covariate.labels = c("% unemployed (logged)", "County college experience (binary)",
                               "% unemployed (logged)*college experience", "Constant"),
          dep.var.labels = "% of county that voted for Trump in 2016",
          dep.var.caption = "",
          column.labels = c("interaction model", "low education", "high education"),
          model.numbers          = FALSE, 
          notes="<span>&#42;</span> p<0.05",
          notes.append=F)
```

